Live and Let Die? (Delayed Issues not Harder to Resolve)

### 2015-08-02 phaseDelay.tex
Updates from ShareLaTeX

% \begin{figure}[!t]
% \scriptsize
% \begin{center}
% \begin{tabular}{l@{~}|l@{~}|r@{~}r@{~}|r@{~}l}
%           &        & \multicolumn{2}{c|}{Time to}  &\\
%           &        & \multicolumn{2}{c|}{resolve} &\\
% Injection & Removal& initial & now & \multicolumn{2}{l}{Scale up w.r.t initial} \\\hline
% Phase $i$&Phase $i$ & 1 & 1 & 1 & ** \\
% Phase $i$ &Phase $i+1$ & 1 & 2& 2 & **** \\
% Phase $i$ &Phase $i+2$ & 1 & 2& 2 & **** \\
% Phase $i$ &Phase $i+3$ & 1 & 2 & 2 & **** \\
%  Phase $i$  &Phase $i+4$ &1 & 2& 2 &**** 
%   \end{tabular}\end{center}
%  \caption{Projects data consistent with
%  the phase delay effect of \fig{curve2}.}\label{fig:curve2a}
% \end{figure}
This paper calls into question all the above claims about the delayed
issue effect. We suggest that the delayed issue effect might have been an dominant
effect decades ago,  but {\em not}  for $21^{\mathit{st}}$
software development. 
 The delayed issue effect was first reported in 1981 in a era of punch card programming
and non-interactive environments~\cite{Boehm81}. In the 21$^\mathit{st}$} century, we  program in 
interactive environments with higher-level languages and better source code control
tools. Such tools allow for the faster refactoring of existing
code-- in which case, 
managing the changes required to fix (say) an incorrect requirements assumption
is far less    onerous    than before.

Also, we note that our development practices have changed in ways that could   mitigate the delayed issued effect.  Previously,
software was large monolithic systems that were write once and maintain
forever. Today, there are more architectures that
support  extensive and faster changes to software. 
For example, consider the MEAN stack preferred for web development by
the continuous deployment community. Older architectures for web development
often used some variant of LAMP (Linux+ Appache+ Mysql+ Php) that was an intricate
combination of tools written in different languages. MEAN stacks, on the other hand,
use Javscript throughout-- which makes large scale reorganizations faster to complete~\cite{wayner15}.
For another example where organizations are adopting increased development flexability, 
Microsoft is adjusting its development practices towards a continuous
release paradigm. Theisen et al. reports their experiences where Microsoft architects
continually adjust  their modules in response to security issues~\cite{theisen15}.
Also,
upgrades to their Windows operating system is moving from service patches (which occur rarely) to continuous deployment (so there will be no Windows 11- just a stream of continuous updates to what is currently called Windows 10~\cite{bright15}).

Other large development shops are similarly reorganizing their work-- witness the US Department of Defense's 2010 mandate
that all new software acquisitions must adopt agile methods~\cite{kim13}. 
This change in DoD policy arises from a separation of baseline architecture
(e.g. the design of an 
aircraft carrier) and the development of applications within that architecture.
For the baseline architecture, bad decisions
made  early in the life cycle may be too expensive to change.
But at least at the DoD, the majority of software development occurs within
the framework of some larger architecture (e.g. an aircraft carrier).
These smaller projects can certainly lever the agile advantage, while 
using fast refactoring tools that mitigate against  the delayed issue effect.

In summary, there exists (1)~historical reasons for believing in the delayed issued
effect as well as (2)~reasons to believe that that effect may not longer hold.
Hence, the rest of this article explores the currency of the delayed issue effect.
After some initial definitions for terminology and hypotheses, we discuss the
value of checking old ideas. Next, we offer a  survey of industrial practitioners and researchers to show that this the negative impact of issue delay is the very prominent
concern in thos communities (in fact, amongst industrial practitioners, it is their number one concern).   Next, we  analyze 171 software  projects developed in the period 2006 to 2014 to find  {\em no evidence} of the delayed issue effect
(note that, to ensure reproducibility,
all the data  used in this study is available in the PROMISE
repository at openscience.us/repo). Finally, we discuss the validity and implications of our results.

\section{Definitions \& Hypotheses}
This paper uses the following definitions:
\bi
\item
{\em Delayed issue effect}:   it is {\em exponentially}  more {\em difficult} to resolve an issue in a software project, the {\em longer} it remains.
\item
 {\em Longer} time is defined as per  Boehm'81~\cite{Boehm81}; i.e. the gap between the   phases where   issues are introduced and resolved.
\item
A measure collected in phase ${1,.,i,..j}$ is 
{\em exponential} when  that
   measure at phase $j$   
   is larger than the sum of that measure in 
earlier phases $1 \le i < j$. 
\item
Issues are more {\em difficult}  
when their resolution takes more time or costs more  (e.g. needs expensive
debugging tools or the skills of expensive developers).
\ei
This paper explores the  following hypotheses.
\subsubsection*{H1: DIE is a current and widely-held belief}

Using a literature review, we can confirm that there are numerous historical papers (dating back decades)
that endorse DIE. Also, using a survey conducted from this paper, we find  that 
  DIE  appears up as the
single most strongly-held beliefs amongst commercial software engineers.

\subsubsection*{H2: DIE is a poorly documented effect}

As discussed in our literature review,  many of the papers reporting the DIE
effect are either:
\bi
\item
Quite old (dating from last century);
\item
Just quote prior papers without presenting   new data; 
\item
Or cite data sources that can no longer be
confirmed.
\ei

\subsubsection*{H3: During development, issues   exhibit a DIE effect}
While this paper  endorse the above hypotheses, we will argue against this
third hypothesis. In our sample of 
 171 software  projects developed, we will show
{\em no evidence}  that, during development, issues were exponentially harder to resolve the longer they were left
 in the software.

@@

Before going any furhter,  we digress to discuss the merits of revisiting
old conclusions in software engineering.

@@

\section{H1: DIE is a current and widely-held belief}
To assess the prevalence of the delayed issue effect,   conducted a survey of software engineers. If our surveyed practitioners make management decisions based on their
understanding of SE theory, then the phase delay effect may well influence their judgment.

@@

\section{H2: DIE is a poorly documented effect}

@@

\section{Data Collection}

@@

\section{Analysis of Phase Delay in 171 TSP Projects}
Our units of analysis are:
\bi
    \item \emph{defects} - individual defects are recorded as line items in the defect logs uploaded to the SEMPR at the SEI. One or more defects are reported against a single \emph{plan item} in the time tracking logs, e.g., a review session, an inspection meeting, a test execution.
    \item \emph{time} - Time is tracked per person per plan item in the time-tracking logs, e.g. a 30 minute design review session involving 3 people will have three time log entries summing to 90 minutes. Time includes the time to analyze, repair, and validate a defect fix.
    \item \emph{time per defect} - The total \# of defects found in a plan item during a removal phase divided by the total time spent on that plan item in that phase.
\ei

@@

\begin{figure}[!t]

@@

\subsection{TSP Project Cost-to-fix curve}
This section discusses our first hypothesis ($H_1$) that   that  the cost-to-resolve issues dramatically increases across phases.

@@

The first thing to note in Figure~\ref{fig:fix-phase-dist}, 
is that this data does not follow the exponential increase feared by Beck's \fig{curve1}:
\bi
\item There is much ``up and down'' in the chart. 
\item If we separate out just the {\em Inspect} or {\em Review} data, we can see that there is some increase in {\em Design} and {\em Code}. However, that effect could be a linear or be just noise. In either case,
it does not look like \fig{curve1}.
\item The time in final testing is particularly low suggesting that few defects survived into testing
(and hence, there were few instances where code required massive rework).
\ei

@@

Given the above evidence, we reject $H_1$: there is no clear evidence that  cost-to-fix of defects dramatically increases across phases.

\subsection{Phase delay analysis}

 \subsubsection{$H_2$: Cost-to-resolve Increases with Delay}

%we calculate phase delay as simply the number of project phases (the set used in our data set is shown in %\fig{waterfall}) during which the defect existed in the system, i.e. the ordinal rank of removal phase minus the %ordinal rank of injection phase. Sub-phases (such as review and inspection) are included in this calculation.
%
\fig{raw} shows the 25th, 50th, 75th percentile
of the time spent on {\em review} or {\em inspect} or {\em test} phases
in our 171 projects.
Note that, in TSP, when developers see issues, they enter {\em review} or 
{\em inspect} or {\em test}
until that issue is retired.

In that figure, the data is split out according to issues that were fixed in phase $Y$ after
being introduced in phase $X$. The data is sub-divided into tables according to $X$;
i.e. according to {\em before, planning, requirements, design} or {\em  code}.

The stars on the right-hand-side of that figure show the scale ups associated with the phase removal stages,
expressed as a ratio of the first item in that block (which, by definition, is 1.00 for
that first item).
If $H_2$ was correct, we would expect those bars in each sub-division to exhibit a  shape
like \fig{curve1}.

They do not. Hence, we reject $H_2$.

 \subsubsection{$H_3$: Requirements Repairs are Most Expensive}

As to $H_3$, the {\em requirements} the  median time spent fixing errors in {\em requirements} was 92,93  49,33,39,49,67,122 minutes\footnote{If this time seems long, then recall
that it includes the time required  to (a)~collect data and realize there is an error;
(b)~prepare a fix;  and (c)~apply some validation
procedure to check the fix (e.g. discuss it with a colleague or execute some tests).}.
This is:
\bi
\item
Clearly longer than median times seen {\em Before } and in {\em Planning}
(30,28,45,77,65,56,24,35,33);
\item
Somewhat shorter, but still not exponentially smaller,  than the median times see after {\em requirements} in {\em design}
(103,101, 79,66, 119,88, 80,45) and in {\em code} (65,115, 113,125, 95,133). 
\ei
For completeness, we ran an  
A12 test comparing the raw data (and not just the medians of \fig{raw}) and found
 a non-small effect only for the difference {\em requirements}
to {\em code}. But even though the effect was non-small, 
 the overall size of the difference
{\em requirements} to {\em code} was less than 30\% (i.e. much less than we would
expect from \fig{curve1}.

Based on the above, we reject   $H_3$.

@@

@@

We would argue that this is not the case.
An  exponential DIE effect can be seen {\em during the development} of all
the papers we have reviewed. For example:
\bi
\item In \fig{b81}: see the data relating to development and acceptance testing;
\item In  all the case studies of \fig{steck04} in the columns {\em Requirements} to {\em Test}.
\ei
That is, in the literature we have surveyed, the post-development DIE effect would have been predicted by data collected during pre-deployment development.
Such

@@

\subsection{Construct validity}
The essential constructs in our analyses are \emph{cost-to-fix, defects, and phase delay}.

@@

\subsection{Reliability}
A previous paper~\cite{shirai14} has applied a range of sanity checks to the data.
A common property of real-world data sets is the presence
of noisy entries (superfluous  or spurious data). 
The level of noise can be quite high. As reported
in \cite{shepperd12}, around
10\% to 30\%
of the records in the NASA MDP defect data sets are
affected by noise. Nichols et al.~\cite{shirai14}  report that
the noise levesl in the SEI TSP data are smaller than those seen
in other data sets. They found in the SEI TSP data that:\bi 
\item
4\% of the data was incorrect such as  null values of illegal formats;
\item  2\% of the data has inconsistencies such as timestamps
where the stop time was before the start time;
\item 3\% of the data had data that was not credible
such as tasks listed in one day that took like than six hours.
\ei

@@

\section*{Acknowledgements}
The authors gratefully acknowledge the careful comments of anonymous reviewers-- particularly, their comments on issues relating to ``cost''
and ``time''.
This work was partially funded by an National Science
Foundation grants NSF-CISE 1302169 and CISE 1506586.
\\
We thank Tuma Solutions for providing the Team Process Data Warehouse software in support of this research.  \\
Personal Software Process$\textsuperscript{SM}$, Team Software Process$\textsuperscript{SM}$, and TSP$\textsuperscript{SM}$ are service marks of Carnegie Mellon University.

>>>>

% \begin{figure}[!t]
% \scriptsize
% \begin{center}
% \begin{tabular}{l@{~}|l@{~}|r@{~}r@{~}|r@{~}l}
%           &        & \multicolumn{2}{c|}{Time to}  &\\
%           &        & \multicolumn{2}{c|}{resolve} &\\
% Injection & Removal& initial & now & \multicolumn{2}{l}{Scale up w.r.t initial} \\\hline
% Phase $i$&Phase $i$ & 1 & 1 & 1 & ** \\
% Phase $i$ &Phase $i+1$ & 1 & 2& 2 & **** \\
% Phase $i$ &Phase $i+2$ & 1 & 2& 2 & **** \\
% Phase $i$ &Phase $i+3$ & 1 & 2 & 2 & **** \\
%  Phase $i$  &Phase $i+4$ &1 & 2& 2 &**** 
%   \end{tabular}\end{center}
%  \caption{Projects data consistent with
%  the phase delay effect of \fig{curve2}.}\label{fig:curve2a}
% \end{figure}
This paper calls into question all the above claims about the delayed
issue effect. We suggest that the delayed issue effect might have been an dominant
effect decades ago,  but {\em not}  for $21^{\mathit{st}}$
software development. 
 The delayed issue effect was first reported in 1981 in a era of punch card programming
and non-interactive environments~\cite{Boehm81}. In the 21$^\mathit{st}$ century, we  program in 
interactive environments with higher-level languages and better source code control
tools. Such tools allow for the faster refactoring of existing
code-- in which case, 
managing the changes required to fix (say) an incorrect requirements assumption
is far less    onerous    than before.

Also, we note that our development practices have changed in ways that could   mitigate the delayed issued effect.  Previously,
software was large monolithic systems that were write once and maintain
forever. Today, there are more architectures that
support  extensive and faster changes to software. 
For example, 
Microsoft is adjusting its development practices towards a continuous
release paradigm. Theisen et al. reports their experiences where Microsoft architects
continually adjust  their modules in response to security issues~\cite{theisen15}.
Also, upgrades to their Windows operating system is moving from service patches (which occur rarely) to continuous deployment (so there will be no Windows 11- just a stream of continuous updates to what is currently called Windows 10~\cite{bright15}).

For another example, consider the MEAN stack preferred for web development by
the continuous deployment community. Older architectures for web development
often used some variant of LAMP (Linux + Appache + Mysql + PHP) that was an intricate
combination of tools written in different languages. MEAN stacks, on the other hand,
use Javascript throughout-- which makes large scale reorganizations faster to complete~\cite{wayner15}.

Inspired by the success of agile approaches, other  development organizations are similarly reorganizing their work-- witness the US Department of Defense's 2010 mandate
that all new software acquisitions must adopt agile methods~\cite{kim13}. 
This change in DoD policy arises from a separation of baseline architecture
(e.g. the design of an 
aircraft carrier) and the development of applications within that architecture.
For the baseline architecture, bad decisions
made  early in the life cycle may be too expensive to change.
But at least at the DoD, the majority of software development occurs within
the framework of some larger architecture (e.g. an aircraft carrier).
These smaller projects can certainly lever the agile advantage, while 
using fast refactoring tools that mitigate against  the delayed issue effect.

In summary, there exists (1)~historical reasons for believing in the delayed issued
effect as well as (2)~reasons to believe that that effect may not longer hold.
Hence,  this article explores the currency of the delayed issue effect.
After some initial definitions, we discuss the
value of checking old ideas. Next, survey of industrial practitioners and researchers to document a widespread belief that delayed issues have a negative impact on projects.  Next, we  analyze 171 software  projects developed in the period 2006 to 2014 to find  {\em no evidence} of the delayed issue effect
(note that, to ensure reproducibility,
all the data  used in this study is available in the PROMISE
repository at openscience.us/repo). Finally, we discuss the validity and implications of our results.

\section{Definitions \& Hypotheses}
This paper uses the following definitions:
\bi
\item
{\em Delayed issue effect}:   it is {\em exponentially}  more {\em difficult} to resolve an issue in a software project, the {\em longer} it remains.
\item
 {\em Longer} time is defined as per  Boehm'81~\cite{Boehm81}; i.e. the gap between the   phases where   issues are introduced and resolved.
\item
A measure collected in phase ${1,.,i,..j}$ is 
{\em exponential} when  that
   measure at phase $j$   
   is larger than the sum of that measure in 
earlier phases $1 \le i < j$. 
\item
Issues are more {\em difficult}  
when their resolution takes more time or costs more  (e.g. needs expensive
debugging tools or the skills of expensive developers).
\ei
Note that we use  the  term ``delayed issue effect'' as generalization of the
more specific rule  ``requirements errors are hardest to fix''.
This generalization is valid since the rationale for the rule about requirements
is usually done as per McConnell~\cite{mcconnell01}; i.e. small upstream mistakes very
early in the system can cascade into huge problems later in the lifecycle.
That said, we prefer our more general term ``delayed issue effect'' since not
only might requirements errors cascade, so too might analysis errors, design errors, etc.

One technical aside: there are several ways to  measure {\em difficulty}. The first is a direct measure of the actual work time spent in correcting the defect in the product. Direct measurement enables analysis of the variability the cost-to-fix among a category of defects. The second approach measures the total time spent in a defect removal phase and divides by the total defects removed. This provides an average cost-to-fix including overhead of that activity. For example, a code inspection may require an hour of effort to remove three defects for a 3 defects per hour or to invert the ratio, 1/3 hours per defect. One observation made by Nichols while studying our 171 projects is that most of the time  is spent locating the defects not actually fixing them. we use the second approach in our analysis.

This paper defects the  following claims. Note that we call them
``claims'' not ``hypotheses'' since the later require defense via some statistical
significance test. On the other hand, our claims will be supported via a variety
of arguments, presented later in this paper.

{\bf  Claim1: ``DIE'' is a  widely-held belief.}
Using a literature review, we can confirm that there are numerous historical papers (dating back decades)
that endorse DIE. Also, using a survey conducted from this paper, we find  that 
  DIE  appears up as the
single most strongly-held beliefs amongst commercial software engineers.

{\bf  Claim2: ``DIE'' is a poorly documented.}
 As discussed in our literature review,  many of the papers reporting the DIE
effect are either (1)~quite old (papers dating from last century);
(2)~that just quote prior papers without presenting   new data; 
(3)~or cite data sources that can no longer be
confirmed.

{\bf Claim3: Delayed Issues not Harder to Resolve.}
 In our sample of 
 171 software  projects developed, we will show
{\em no evidence}  that, during development, delayed issues were exponentially harder to resolve the longer they were left
 in the software.

@@

Before going any further,  we digress to discuss the merits of revisiting
old conclusions in software engineering.

@@

\section{``DIE'' is a  widely-held belief}
To assess the prevalence of the delayed issue effect,   conducted a survey of software engineers. If our surveyed practitioners make management decisions based on their
understanding of SE theory, then the phase delay effect may well influence their judgment.

@@

\section{``DIE'' is Poorly  Documented}

@@

\section{Delayed Issues not Harder\newline to Resolve}

@@

Using the above,
our units of analysis are:
\bi
    \item \emph{defects} - individual defects are recorded as line items in the defect logs uploaded to the SEMPR at the SEI. One or more defects are reported against a single \emph{plan item} in the time tracking logs, e.g., a review session, an inspection meeting, a test execution.
    \item \emph{time} - Time is tracked per person per plan item in the time-tracking logs, e.g. a 30 minute design review session involving 3 people will have three time log entries summing to 90 minutes. Time includes the time to analyze, repair, and validate a defect fix.
    \item \emph{time per defect} - The total \# of defects found in a plan item during a removal phase divided by the total time spent on that plan item in that phase.
\ei

@@

\begin{figure}[!b]

@@

\subsection{Observations from 171 Projects}

@@

\fig{raw} shows the 25th, 50th, 75th percentile
of the time spent on {\em review} or {\em inspect} or {\em test} phases
in our 171 projects.
Note that, in TSP, when developers see issues, they enter {\em review} or 
{\em inspect} or {\em test}
until that issue is retired.

In that figure, the data is split out according to issues that were fixed in phase $Y$ after
being introduced in phase $X$. The data is sub-divided into tables according to $X$;
i.e. according to {\em before, planning, requirements, design} or {\em  code}. 
The stars on the right-hand-side of that figure show the scale ups associated with the phase removal stages,
expressed as a ratio of the first item in that block (which, by definition, is 1.00 for
that first item).
If  this data exhibited a delayed issue effect,  we would expect those bars in each sub-division to exhibit an
exponential shape.

Using Figure~\ref{fig:fix-phase-dist}, we can comment on two specific and more general
delayed issued effects. 
\bi
\item {\bf Specific1}: Requirements errors are the most expensive to fix;
\item {\bf Specific2}: Requirements errors are the most expensive to fix, 
 when found during production but the cheapest to fix early in development;
 \item {\bf General:}  It is exponentially more  difficult  to resolve an issue in a software project, the   longer it remains.
\ei
\subsubsection{Specific1}

Our data does not support   
{\em  requirements errors are the most expensive to fix}.
Note that   median time spent fixing errors from {\em requirements} across the life cycle was 92, 93,  49, 33, 39, 49, 67, 122 minutes\footnote{If this time seems long, then recall
that it includes the time required  to (a)~collect data and realize there is an error;
(b)~prepare a fix;  and (c)~apply some validation
procedure to check the fix (e.g. discuss it with a colleague or execute some tests).}.
This is:
\bi
\item
Clearly longer than median times seen {\em Before } and in {\em Planning}
(30,28,45,77,65,56,24,35,33);
\item
Somewhat shorter, but still not exponentially smaller,  than the median times see after {\em requirements} in {\em design}
(103,101, 79,66, 119,88, 80,45) and in {\em code} (65,115, 113,125, 95,133). 
\ei
For completeness, we ran an  
A12 test comparing the raw data (and not just the medians of \fig{raw}) and found
 a non-small effect only for the difference {\em requirements}
to {\em code}. But even though the effect was non-small, 
 the overall size of the difference
{\em requirements} to {\em code} was less than 30\% (i.e. much less than we would
expect from the delayed issue effect). 
\subsubsection{Specific2}
Next, we comment on 
requirements errors are the most expensive to fix 
{\em when found during production but the cheapest to fix early in development}}. 
Our data from these 171 project does not extend into post-delivery deployment. That said,
there is enough evidence to raise doubts that our projects will exhibit exponentially more
difficult issues after deployment. 
Specifically: prior reports of an  exponential increase in difficulty in resolving issues {\em during
deployment} might have been used to predict for exponentially increasing
difficulty {\em post-deployment}.
For example:
\bi
\item  From requirements to acceptance testing
\fig{b81} reports a 40-fold increase;
\item
 \fig{cost-to-fix} reports a 100-fold increase
(for the larger projects).
\item
The projects surveyed in \fig{steck} report changes of a  \newline
\{37,40,40,50,51,70,75,250\}-fold increase in pre-deployment period when the software was being developed.
\ei
On the other hand, a remarkable feature of Figure~\ref{fig:fix-phase-dist} is the {\em absence}
of an exponential delay effect in the difficulty to resolve requirements issues during development.
The greatest difficulty in resolving requirements issues was 1.31 seen in systems testing--
which is far smaller than the 37 to 250-fold increases reported in \fig{b81}, \fig{cost-to-fix},
and \fig{steck}.  That is, from our TSP data, we would not anticipate exponentially greater difficulty in resolving requirements issues after
deployment. 
 \subsubsection{General}
 The above specific statements related just to requirements. Turning now
 to issues creased in all phases in Figure~\ref{fig:fix-phase-dist},  we see that
 this data does not follow the exponential increase predicted
by the delayed issue effect:
\bi
\item There is much ``up and down'' in the chart. 
\item If we separate out just the {\em Inspect} or {\em Review} data, we can see that there is some increase in {\em Design} and {\em Code}. However, that effect could be a linear or be just noise. In either case,
it does not look like exponential.
\item The time in final testing is particularly low suggesting that few defects survived into testing
(and hence, there were few instances where code required massive rework).
\ei

@@

%we calculate phase delay as simply the number of project phases (the set used in our data set is shown in %\fig{waterfall}) during which the defect existed in the system, i.e. the ordinal rank of removal phase minus the %ordinal rank of injection phase. Sub-phases (such as review and inspection) are included in this calculation.
%

@@

To sum up the paper so far,
\bi
\item
The delayed issue effect has been reported in around a dozen papers, at least half of which date back to the 1970s. 
\item
But there exists
no evidence for this effect was seen  in hundreds of software projects build   since 2006.
\ei
From this second point, is it valid to assert that it is {\em not} disruptive to  software projects
to let some issues advance from one phase to the next?

\subsection{Alternate Explanations}

One explanation for the above is that our sample is biased-- that while general software
systems suffer from the delayed issue effect, systems built using TSP are so much better controlled
that they  mitigate against the DIE effect.

It would be convenient to offer the last paragraph as the conclusion of this study. Two of our authors (Nichols and Shull)
 work at the organization that invented and promoted TSP.

@@

We would argue that this is not the case.
An  exponential DIE effect can be seen {\em during the development} of all
the papers we have reviewed. For example:
\bi
\item In \fig{b81}: see the data relating to development and acceptance testing;
\item In  all the case studies of \fig{steck} in the columns {\em Requirements} to {\em Test}.
\ei
That is, in the literature we have surveyed, the post-development DIE effect would have been predicted by data collected during pre-deployment development.
Such

@@

\subsection{Construct validity}
XXXX die does not aoccur during developmentjust during deployment

that means during period requirements to test pre-deployent, there is not rush to resolve issues (just as long as a concerted effort is made to retire all issues priot to deployment)

all our data is skewed due to TSP. two responses. 1) we don't thing so. 2) sure, let it due to TSP in which case we can fix the delayed issue effect via process methods 9everyone adopts the team software process). In either case, the delayed issue effect can be ignored or mitigated without extensive investment in radically new methods or tools (e.g. automatic formal methods, better requirement description languages).

The essential constructs in our analyses are \emph{cost-to-fix, defects, and phase delay}.

@@

@@

\section*{Acknowledgements}
The authors wish to thank  David Tuma and  Yasutaka Shirai for their work on the SEI databases
that made this analysis possible.
In particular, we thank Tuma Solutions for providing the Team Process Data Warehouse software.
Also, the authors gratefully acknowledge the careful comments of anonymous reviewers-- particularly, their comments on issues relating to ``cost''
and ``time''.
This work was partially funded by an National Science
Foundation grants NSF-CISE 1302169 and CISE 1506586.
Personal Software Process$\textsuperscript{SM}$, Team Software Process$\textsuperscript{SM}$, and TSP$\textsuperscript{SM}$ are service marks of Carnegie Mellon University.
